The Proofs

### 2024-07-16 proofs/tex/norm-lemmas.tex
WIP: fix references, add appendix

      $[[ [nf(σ)] nf(iN) ]] = [[ [nf(σ)] nf(na) ]] = [[ [nf(σ)] na ]] $.
    \item $[[iP]]$ = $[[pa]]$ \\
      Similar to \cref{case:norm-subst-distr-var}.
   \item If the type is formed by $[[→]]$, $[[↑]]$, or $[[↓]]$, 
     the required equality follows from the congruence of the normalization and
     substitution and the induction hypothesis.
     For example, if $[[iN]] = [[iP → iM]]$ then \\
     $\begin{aligned}[t]
        [[nf([σ] iN)]] &= [[ nf([σ] (iP → iM)) ]] \\
                        &= [[ nf([σ]iP → [σ]iM) ]]
                        && \text{By congruence of substitution} \\
                        &= [[ nf([σ]iP) → nf([σ]iM) ]]
                        && \text{By congruence of normalization} \\
                        &= [[ [nf(σ)]nf(iP) → [nf(σ)]nf(iM) ]]
                        && \text{By induction hypothesis} \\
                        &= [[ [nf(σ)](nf(iP) → nf(iM)) ]]
                        && \text{By congruence of substitution} \\
                        &= [[ [nf(σ)]nf(iP → iM) ]]
                        && \text{By congruence of normalization} \\
                        &= [[ [nf(σ)]nf(iN) ]]
      \end{aligned}$ \\
    \item $[[iN]] = [[∀ pas.iM]]$ \label{case:norm-subst-commute} \\
      $\begin{aligned}[t]
          [[ [nf(σ)] nf(iN) ]] &= [[ [nf(σ)] nf(∀pas.iM)]] \\
                            &= [[ [nf(σ)] ∀pas'.nf(iM) ]]
                            && \text {Where $[[pas']] = [[ ord {pas} in nf(iM)]]
                               = [[ord {pas} in iM]]$}\\
                           & && 
                               \text{(the latter is by
                               \cref{corollary:normalization-ord})}\\
        \end{aligned}$ \\

      $\begin{aligned}[t]
         [[ nf([σ]iN) ]] &= [[ nf([σ] ∀pas.iM)]] \\
                          &= [[ nf(∀pas.[σ]iM) ]]
                          && \text{Assuming $[[{pas} ∩ Γ1]] = \emptyset$
                             and $[[{pas} ∩ Γ2]] = \emptyset$}\\
                          &= [[ ∀pbs.nf([σ]iM) ]]
                          && \text {Where $[[pbs]] = [[ord {pas} in nf([σ]iM)]]
                             = [[ord {pas} in [σ]iM]]$}\\
                          & && 
                             \text{(the latter is by \cref{corollary:normalization-ord})}\\
                          &= [[ ∀pas'.nf([σ]iM) ]]
                          && \text{By \cref{lemma:ord-sigma},}\\
                          & && \text{$[[pbs]] = [[pas']]$
                             since $[[{pas}]]$ is disjoint with $[[Γ1]]$ and
                             $[[Γ2]]$}\\
                          &= [[ ∀pas'.[nf(σ)]nf(iM) ]]
                          && \text {By the induction hypothesis}\\
         \end{aligned}$ \\

     To show the alpha-equivalence of 
     $[[ [nf(σ)] ∀pas'.nf(iM) ]]$ and $[[ ∀pas'.[nf(σ)]nf(iM) ]]$,
     we can assume that $[[{pas'} ∩ Γ1]] = \emptyset$, and $[[{pas'} ∩ Γ2]]
     = \emptyset$.

@@

\lemmaNormalizationCompleteness*
\begin{proof}
  Mutual induction on $[[iN ≈ iM]]$ and $[[iP ≈ iQ]]$.
  \begin{caseof}
  \item {\nameref{\ottdruleEOneForallLabel}} \label{case:ord-completeness:forall} \\

    From the definition of the normalization,
    \begin{itemize}
      \item $[[nf(∀pas.iN)]] = [[∀pas'.nf(iN)]]$ where $[[pas']]$ is $[[ord {pas} in nf(iN)]]$
      \item $[[nf(∀pbs.iM)]] = [[∀pbs'.nf(iM)]]$ where $[[pbs']]$ is $[[ord {pbs} in nf(iM)]]$
    \end{itemize}
    Let us take $[[mu : ({pbs} ∩ fv iM) ↔ ({pas} ∩ fv iN)]]$ from the
    inversion of the equivalence judgment. Notice that from
    \cref{lemma:fv-nf,lemma:ord-soundness}, the domain and the codomain of $\mu$ can be written
    as $[[mu : {pbs'} ↔ {pas'}]]$.

@@

    \item[(ii)] $\begin{aligned}[t] [[ [mu]pbs' ]] &= [[ [mu] ord {pbs} in nf(iM) ]]
                                                  && \text{by the definition of $[[pbs']]$ } \\
                                                  &= [[ [mu] ord ({pbs} ∩ fv iM) in nf(iM) ]]
                                                  && \text{from \cref{lemma:fv-nf,corollary:ord-weakening} } \\
                                                  &= [[ ord [mu] ({pbs} ∩ fv iM) in [mu] nf(iM) ]]
                                                  && \text{by \cref{lemma:distr-mu-ord}, because}\\
                                                  & && \text{$[[{pas} ∩ fv iN]] \cap [[fv nf(iM)]] \subseteq [[{pas} ∩ fv iM ]]
                                                     = \emptyset$}\\
                                                  &
                                                  && \text{$[[{pas} ∩ fv iN]] \cap [[({pbs} ∩ fv iM)]] \subseteq
                                                     [[{pas} ∩ fv iM]] = \emptyset$} \\
                                                  &= [[ ord [mu] ({pbs} ∩ fv iM) in nf(iN) ]]
                                                  && \text{since $[[ [mu] nf(iM) ]] = [[nf(iN)]]$ is proved } \\
                                                  &= [[ ord ({pas} ∩ fv iN) in nf(iN) ]]
                                                  && \text{because $\mu$ is a bijection between}\\
                                                  & && \text{$[[{pas} ∩ fv iN]]$ and $[[{pbs} ∩ fv iM]]$} \\
                                                  &= [[ ord {pas} in nf(iN) ]]
                                                  && \text{from \cref{lemma:fv-nf,corollary:ord-weakening} } \\
                                                  &= [[ pas' ]]
                                                  && \text{by the definition of $[[pas']]$} \\
      \end{aligned}$
    \end{enumerate}
  \item {\nameref{\ottdruleEOneExistsLabel}} Same as for \cref{case:ord-completeness:forall}.
  \item Other rules are congruent, and thus, proved by the corresponding congruent alpha-equivalence rule,
    which is applicable by the induction hypothesis. 
  \end{caseof}
\end{proof}

@@

  \item $[[iN]] = [[iP → iM]]$\\
    \label{case:normal-after-subst-arrow}
    $
    \begin{aligned}[t]

@@

    $

@@

  \item $[[iN]] = [[∀pas.iM]]$\\
    $
    \begin{aligned}[t]

@@

\corSubstPresDeclEquiv*
\begin{proof}
  $
  \begin{aligned}[t] 
    [[ iP ≈ iQ ]] &\Rightarrow        [[ nf(iP) = nf(iQ) ]]
                  && \text{by \cref{lemma:subt-equiv-algorithmization}}\\
                  &\Rightarrow [[ [nf(σ)]nf(iP) = [nf(σ)]nf(iQ)]]\\
                  &\Rightarrow [[ nf([σ]iP) = nf([σ]iQ)]]
                  && \text{by \cref{lemma:norm-subst-distr}}\\ 
                  &\Rightarrow        [[ [σ]iP ≈ [σ]iQ ]]
                  && \text{by \cref{lemma:subt-equiv-algorithmization}}\\
  \end{aligned} 
  $
\end{proof}

>>>>

      $[[ [nf(σ)] nf(iN) ]] = [[ [nf(σ)] nf(na) ]] = [[ [nf(σ)] na ]] $.
    \item $[[iP]]$ = $[[pa]]$ \\
      Similar to \cref{case:norm-subst-distr-var}.
   \item If the type is formed by $[[→]]$, $[[↑]]$, or $[[↓]]$, 
     the required equality follows from the congruence of the normalization and
     substitution and the induction hypothesis.
     For example, if $[[iN]] = [[iP → iM]]$ then
     \begin{align*}
        [[nf([σ] iN)]] &= [[ nf([σ] (iP → iM)) ]] \\
                        &= [[ nf([σ]iP → [σ]iM) ]]
                        && \text{By congruence of substitution} \\
                        &= [[ nf([σ]iP) → nf([σ]iM) ]]
                        && \text{By congruence of normalization} \\
                        &= [[ [nf(σ)]nf(iP) → [nf(σ)]nf(iM) ]]
                        && \text{By induction hypothesis} \\
                        &= [[ [nf(σ)](nf(iP) → nf(iM)) ]]
                        && \text{By congruence of substitution} \\
                        &= [[ [nf(σ)]nf(iP → iM) ]]
                        && \text{By congruence of normalization} \\
                        &= [[ [nf(σ)]nf(iN) ]]
      \end{align*}
    \item $[[iN]] = [[∀ pas.iM]]$ \label{case:norm-subst-commute}
      \begin{align*}
          [[ [nf(σ)] nf(iN) ]] &= [[ [nf(σ)] nf(∀pas.iM)]] \\
                            &= [[ [nf(σ)] ∀pas'.nf(iM) ]]
                            && \text {Where $[[pas']] = [[ ord {pas} in nf(iM)]]
                               = [[ord {pas} in iM]]$}\\
                           & && 
                               \text{(the latter is by
                               \cref{corollary:normalization-ord})}\\
        \end{align*}

      \begin{align*}
         [[ nf([σ]iN) ]] &= [[ nf([σ] ∀pas.iM)]] \\
                          &= [[ nf(∀pas.[σ]iM) ]]
                          && \text{Assuming $[[{pas} ∩ Γ1]] = \emptyset$
                             and $[[{pas} ∩ Γ2]] = \emptyset$}\\
                          &= [[ ∀pbs.nf([σ]iM) ]]
                          && \text {Where $[[pbs]] = [[ord {pas} in nf([σ]iM)]]
                             = [[ord {pas} in [σ]iM]]$}\\
                          & && 
                             \text{(the latter is by \cref{corollary:normalization-ord})}\\
                          &= [[ ∀pas'.nf([σ]iM) ]]
                          && \text{By \cref{lemma:ord-sigma},}\\
                          & && \text{$[[pbs]] = [[pas']]$
                             since $[[{pas}]]$ is disjoint with $[[Γ1]]$ and
                             $[[Γ2]]$}\\
                          &= [[ ∀pas'.[nf(σ)]nf(iM) ]]
                          && \text {By the induction hypothesis}\\
         \end{align*}
     To show the alpha-equivalence of 
     $[[ [nf(σ)] ∀pas'.nf(iM) ]]$ and $[[ ∀pas'.[nf(σ)]nf(iM) ]]$,
     we can assume that $[[{pas'} ∩ Γ1]] = \emptyset$, and $[[{pas'} ∩ Γ2]]
     = \emptyset$.

@@

\lemmaNormalizationCompleteness*
\begin{proof}
  Mutual induction on $[[iN ≈ iM]]$ and $[[iP ≈ iQ]]$.
  \begin{caseof}
  \item {\nameref{\ottdruleEOneForallLabel}} \label{case:ord-completeness:forall}
   From the normalization definition,
    \begin{itemize}
      \item $[[nf(∀pas.iN)]] = [[∀pas'.nf(iN)]]$ where $[[pas']]$ is $[[ord {pas} in nf(iN)]]$
      \item $[[nf(∀pbs.iM)]] = [[∀pbs'.nf(iM)]]$ where $[[pbs']]$ is $[[ord {pbs} in nf(iM)]]$
    \end{itemize}
    Let us take $[[mu : ({pbs} ∩ fv iM) ↔ ({pas} ∩ fv iN)]]$ from the
    inversion of the equivalence judgment. Notice that from
    \cref{lemma:fv-nf,lemma:ord-soundness}, the domain and the codomain of $\mu$ can be written
    as $[[mu : {pbs'} ↔ {pas'}]]$.

@@

    \item[(ii)] 
    \begin{align*} 
      [[ [mu]pbs' ]] &= [[ [mu] ord {pbs} in nf(iM) ]]
                      && \text{by the definition of $[[pbs']]$ } \\
                      &= [[ [mu] ord ({pbs} ∩ fv iM) in nf(iM) ]]
                      && \text{from \cref{lemma:fv-nf,corollary:ord-weakening} } \\
                      &= [[ ord [mu] ({pbs} ∩ fv iM) in [mu] nf(iM) ]]
                      && \text{by \cref{lemma:distr-mu-ord}, because}\\
                      & && \text{$[[{pas} ∩ fv iN]] \cap [[fv nf(iM)]] \subseteq [[{pas} ∩ fv iM ]]
                        = \emptyset$}\\
                      &
                      && \text{$[[{pas} ∩ fv iN]] \cap [[({pbs} ∩ fv iM)]] \subseteq
                        [[{pas} ∩ fv iM]] = \emptyset$} \\
                      &= [[ ord [mu] ({pbs} ∩ fv iM) in nf(iN) ]]
                      && \text{since $[[ [mu] nf(iM) ]] = [[nf(iN)]]$ is proved } \\
                      &= [[ ord ({pas} ∩ fv iN) in nf(iN) ]]
                      && \text{because $\mu$ is a bijection between}\\
                      & && \text{$[[{pas} ∩ fv iN]]$ and $[[{pbs} ∩ fv iM]]$} \\
                      &= [[ ord {pas} in nf(iN) ]]
                      && \text{from \cref{lemma:fv-nf,corollary:ord-weakening} } \\
                      &= [[ pas' ]]
                      && \text{by the definition of $[[pas']]$} \\
      \end{align*}
    \end{enumerate}
  \item {\nameref{\ottdruleEOneExistsLabel}} Same as for \cref{case:ord-completeness:forall}.
  \item Other rules are congruent, and thus, proved by the corresponding congruent alpha-equivalence rule,
    which is applicable by the induction hypothesis. 
  \end{caseof}
\end{proof}

@@

  \item $[[iN]] = [[iP → iM]]$ \label{case:normal-after-subst-arrow}
    \begin{align*}

@@

    \end{align*}

@@

  \item $[[iN]] = [[∀pas.iM]]$
    \begin{align*}

@@

\corSubstPresDeclEquiv*
\begin{proof}
  \begin{align*} 
    [[ iP ≈ iQ ]] &\Rightarrow        [[ nf(iP) = nf(iQ) ]]
                  && \text{by \cref{lemma:subt-equiv-algorithmization}}\\
                  &\Rightarrow [[ [nf(σ)]nf(iP) = [nf(σ)]nf(iQ)]]\\
                  &\Rightarrow [[ nf([σ]iP) = nf([σ]iQ)]]
                  && \text{by \cref{lemma:norm-subst-distr}}\\ 
                  &\Rightarrow        [[ [σ]iP ≈ [σ]iQ ]]
                  && \text{by \cref{lemma:subt-equiv-algorithmization}}\\
  \end{align*} 
\end{proof}

### 2024-07-16 proofs/tex/ord-lemmas.tex
WIP: fix references, add appendix

      On the other hand, 
      As a set, $[[{ordVars1} ∪ {ordVars2}]]$
      is equal to $[[ordVars1, (ordVars2 \ {ordVars1})]]$. 
    \item  \ruleref{\ottdruleOForallLabel}.
      Then  the inferred ordering judgement has shape
      $[[ord varset in ∀pas.iN = ordVars]]$,
      and by inversion, 
      $[[varset ∩ {pas} = ∅]]$    
      $[[ord varset in iN = ordVars]]$.
      The latter implies that $[[varset ∩ fv iN = {ordVars}]]$.
      We need to show that $[[varset ∩ fv ∀pas.iN = {ordVars}]]$,
      or equivalently, that
      $[[varset ∩ (fv iN \ {pas}) = varset ∩ fv iN ]]$,
      which holds since $[[varset ∩ {pas} = ∅]]$.
  \end{caseof}
\end{proof}

@@

\lemDistrMuOrd*
\begin{proof}
  Mutual induction on $[[iN]]$ and $[[iP]]$.
  \begin{caseof}
  \item $[[iN]]$ = $[[na]]$ \label{case:distr-mu-ord:var} \\
    let us consider four cases:
    \begin{caseof}
    \item $[[na]] \in A$ and $[[na]] \in [[varset]]$\\ Then
    \\
      $
      \begin{aligned}[t] [[ [mu] (ord varset in iN) ]] &= [[ [mu] (ord varset in na)]] \\
                                                             &= [[ [mu] na ]]
                                                             && \text{by \ruleref{\ottdruleOPVarInLabel}}\\
                                                             &= [[nb]]
                                                             && \text{for some $[[nb]] \in B$ (notice $[[nb]] \in [[ [mu]varset ]]$)} \\
                                                             &= [[ ord [mu]varset in nb ]]
                                                             && \text{by \ruleref{\ottdruleOPVarInLabel},
                                                                because $[[nb]] \in [[ [mu]varset ]]$} \\
                                                             &= [[ord [mu] varset in [mu] na ]]
       \end{aligned}
       $
     \item $[[na]] \notin A$ and $[[na]] \notin [[varset]]$\\
       Notice that
       $[[ [mu] (ord varset in iN) ]] = [[ [mu] (ord varset in na)]] = [[·]]$ by
       \ruleref{\ottdruleOPVarNInLabel}.
       On the other hand, $[[ ord [mu] varset in [mu] na = ord [mu] varset
       in na ]] = [[·]]$ The latter equality is from
       \ruleref{\ottdruleOPVarNInLabel}, because
       $[[mu]]$ is collision-free on $[[varset]]$ and $[[fv iN]]$, so
       $[[fv iN]] \ni [[na]] \notin [[mu]](A \cap [[varset]]) \cup
       [[varset]] \supseteq [[ [mu] varset ]]$.
     \item $[[na]] \in A$ but $[[na]] \notin [[varset]]$\\ Then
       $[[ [mu] (ord varset in iN) ]] = [[ [mu] (ord varset in na)]] = [[·]]$
       by \ruleref{\ottdruleOPVarNInLabel}.
       To prove that\\ $[[ ord [mu] varset in [mu] na ]] = [[·]]$, we apply
       \ruleref{\ottdruleOPVarNInLabel}. Let us show that
       $[[ [mu] na ]] \notin [[ [mu] varset ]]$.
       Since $[[ [mu] na ]] = [[mu]]([[na]])$ and
       $[[ [mu] varset ]] \subseteq [[mu]](A \cap [[varset]]) \cup [[varset]]$,
       it suffices to prove 
       $[[mu]]([[na]]) \notin [[mu]](A \cap [[varset]]) \cup [[varset]]$.

@@

  \item $[[iN]] = [[↑iP]]$ \\
    $\begin{aligned}[t]
       [[ [mu] (ord varset in iN) ]] &= [[ [mu] (ord varset in ↑iP) ]] \\
                                     &= [[ [mu] (ord varset in iP) ]]
                                     && \text{by \ruleref{\ottdruleOShiftULabel}}\\
                                     &= [[ ord [mu]varset in [mu]iP ]]
                                     && \text{by the induction hypothesis}\\
                                     &= [[ ord [mu]varset in  ↑[mu]iP ]]
                                     && \text{by \ruleref{\ottdruleOShiftULabel}}\\
                                     &= [[ ord [mu]varset in  [mu]↑iP ]]
                                     && \text{by the definition of substitution}\\
                                     &= [[ ord [mu]varset in  [mu]iN ]]
            \end{aligned}$

   \item $[[iN]] = [[iP → iM]]$  \\
     $\begin{aligned}[t]
        & [[ [mu] (ord varset in iN) ]] \\ 
        &= [[ [mu] (ord varset in iP → iM) ]] \\
        &= [[ [mu] (ordVars1, (ordVars2 \ {ordVars1})) ]]
          && \text{where } [[ord varset in iP = ordVars1]] \text{ and } [[ord varset in iM = ordVars2]] \\
        &= [[ [mu] ordVars1, [mu](ordVars2 \ {ordVars1}) ]] \\
        &= [[ [mu] ordVars1, ([mu]ordVars2 \ [mu]{ordVars1}) ]]
          && \text{by induction on $[[ordVars2]]$;
                  the inductive step is similar to 
                  \cref{case:distr-mu-ord:var}.}\\
        & && \text{Notice that $[[mu]]$ is collision free on $[[{ordVars1}]]$ and $[[{ordVars2}]]$} \\
        & && \text{since
          $[[{ordVars1}]] \subseteq [[varset]]$ and
          $[[{ordVars2}]] \subseteq [[fv iN]]$ }\\
          &= [[ [mu] ordVars1, ([mu]ordVars2 \ {[mu]ordVars1}) ]]
      \end{aligned}$ \\
      \hfill\\
      On the other hand, \\
    $\begin{aligned}[t]
       & [[  ord [mu] varset in [mu]iN ]] \\ 
       &= [[ ord [mu] varset in [mu]iP → [mu]iM ]] \\
       &= [[ ordVarsb1, (ordVarsb2 \ {ordVarsb1}) ]]
       && \text{where } [[ord [mu] varset in [mu] iP = ordVarsb1]] \text{ and }  \\
       & && [[ord [mu] varset in [mu] iM = ordVarsb2]]; 
            \text{ then by the induction}\\
       & && \text{hypothesis, $[[ordVarsb1]] = [[ [mu] ordVars1 ]]$,
                   $[[ordVarsb2]] = [[ [mu] ordVars2 ]]$}\\
       &= [[ [mu] ordVars1, ([mu]ordVars2 \ {[mu]ordVars1}) ]]
     \end{aligned}$

   \item $[[iN]] = [[∀ pas.iM]]$ \\
     $\begin{aligned}[t]
          [[ [mu] (ord varset in iN) ]] &= [[ [mu] ord varset in ∀pas.iM]] \\
                                        &= [[ [mu] ord varset in iM]] \\
                                        &= [[ ord [mu] varset in [mu] iM]]
                                        && \text {by the induction hypothesis}\\
     \end{aligned}$ \\
     $ 
     \begin{aligned}[t]
       [[ (ord [mu] varset in [mu] iN) ]] &= [[ ord [mu] varset in [mu] ∀pas.iM ]] \\
                                          &= [[ ord [mu] varset in ∀pas.[mu]iM ]] \\
                                          &= [[ ord [mu] varset in [mu] iM ]] \\
     \end{aligned}
     $

  \end{caseof}
\end{proof}

\lemOrdSigma*
\begin{proof}
  Mutual induction on $[[iN]]$ and $[[iP]]$.
  \begin{caseof}
    \item $[[iN = na]]$ \\
      If $[[na ∉ Γ1]]$ then $[[ [σ]na = na ]]$ and $[[ ord varset in [σ]na ]] = [[ ord varset in na ]]$, 
      as requried.
      If $[[na ∊ Γ1]]$ then $[[na ∉ varset]]$, so $[[ ord varset in na ]] = [[·]]$.
      Moreover, $[[Γ2 ⊢ σ : Γ1]]$ means $[[ fv([σ]na) ⊆ Γ2 ]]$, and thus, 
      as a set, $[[ ord varset in [σ]na ]] = [[varset ∩ fv([σ]na)]] \subseteq [[varset ∩ Γ2]] = [[·]]$.
    \item $[[iN = ∀pas.iM]]$\\
      We can assume $[[{pas} ∩ Γ1 = ∅]]$
      and $[[{pas} ∩ varset = ∅]]$. Then

      $\begin{aligned}[t]
         [[ ord varset in [σ]iN ]] &= [[ ord varset in [σ]∀pas.iM ]] \\
                                   &= [[ ord varset in ∀pas.[σ]iM ]]\\
                                   &= [[ ord varset in [σ]iM ]]
                                   && \text{by the induction hypothesis}\\
                                   &= [[ ord varset in iM ]]\\
                                   &= [[ ord varset in ∀pas.iM ]]\\
                                   &= [[ ord varset in iN ]]
       \end{aligned}$
    \item $[[iN = ↑iP]]$\\
       $\begin{aligned}[t]
        [[ ord varset in [σ]iN ]] &= [[ ord varset in [σ]↑iP ]] \\
                                   &= [[ ord varset in ↑[σ]iP ]]
                                   && \text{by the definition of substitution}\\
                                   &= [[ ord varset in [σ]iP ]]
                                   && \text{by the induction hypothesis}\\
                                   &= [[ ord varset in iP ]]
                                   && \text{by the definition of substitution}\\
                                   &= [[ ord varset in ↑iP ]]
                                   && \text{by the definition of ordering}\\
                                   &= [[ ord varset in iN ]]
       \end{aligned} $

    \item $[[iN = iP → iM]]$\\
       $ \begin{aligned}
        [[ ord varset in [σ]iN ]] &= [[ ord varset in [σ](iP → iM) ]] \\
                                   &= [[ ord varset in ([σ]iP → [σ]iM) ]]
                                   && \text{def. of substitution}\\
                                   &= [[ ord varset in [σ]iP]],\\
                                   &\phantom{=} ~ [[(ord varset in [σ]iM \ {ord varset in [σ]iP}) ]]
                                   && \text{def. of ordering}\\
                                   &= [[ ord varset in iP]],\\
                                   &\phantom{=} ~ [[(ord varset in iM \ {ord varset in iP}) ]]
                                   && \text{the induction hypothesis}\\
                                   &= [[ ord varset in iP → iM ]]
                                   && \text{def. of ordering}\\
                                   &= [[ ord varset in iN ]]
       \end{aligned} $
    \item The proofs of the positive cases are symmetric.
  \end{caseof}
\end{proof}

>>>>

      On the other hand, 
      as a set $[[{ordVars1} ∪ {ordVars2}]]$
      is equal to $[[ordVars1, (ordVars2 \ {ordVars1})]]$. 
    \item  \ruleref{\ottdruleOForallLabel}.
      Then  the inferred ordering judgement has shape
      $[[ord varset in ∀pas.iN = ordVars]]$,
      and by inversion, 
      $[[varset ∩ {pas} = ∅]]$    
      $[[ord varset in iN = ordVars]]$.
      The latter implies that $[[varset ∩ fv iN = {ordVars}]]$.
      We need to show that $[[varset ∩ fv ∀pas.iN = {ordVars}]]$,
      or equivalently, that
      $[[varset ∩ (fv iN \ {pas}) = varset ∩ fv iN ]]$,
      which holds since $[[varset ∩ {pas} = ∅]]$.
  \end{caseof}
\end{proof}

@@

\lemDistrMuOrd*
\begin{proof}
  Mutual induction on $[[iN]]$ and $[[iP]]$.
  \begin{caseof}
  \item $[[iN]]$ = $[[na]]$ \label{case:distr-mu-ord:var} \\
    let us consider four cases:
    \begin{caseof}
    \item $[[na]] \in A$ and $[[na]] \in [[varset]]$. Then
      \begin{align*} [[ [mu] (ord varset in iN) ]] &= [[ [mu] (ord varset in na)]] \\
                                                             &= [[ [mu] na ]]
                                                             && \text{by \ruleref{\ottdruleOPVarInLabel}}\\
                                                             &= [[nb]]
                                                             && \text{for some $[[nb]] \in B$ (notice $[[nb]] \in [[ [mu]varset ]]$)} \\
                                                             &= [[ ord [mu]varset in nb ]]
                                                             && \text{by \ruleref{\ottdruleOPVarInLabel},
                                                                because $[[nb]] \in [[ [mu]varset ]]$} \\
                                                             &= [[ord [mu] varset in [mu] na ]]
       \end{align*}
     \item $[[na]] \notin A$ and $[[na]] \notin [[varset]]$\\
       Notice that
       $[[ [mu] (ord varset in iN) ]] = [[ [mu] (ord varset in na)]] = [[·]]$ by
       \ruleref{\ottdruleOPVarNInLabel}.
       On the other hand, $[[ ord [mu] varset in [mu] na = ord [mu] varset
       in na ]] = [[·]]$ The latter equality is from
       \ruleref{\ottdruleOPVarNInLabel}, because
       $[[mu]]$ is collision-free on $[[varset]]$ and $[[fv iN]]$, so
       $[[fv iN]] \ni [[na]] \notin [[mu]](A \cap [[varset]]) \cup
       [[varset]] \supseteq [[ [mu] varset ]]$.
     \item $[[na]] \in A$ but $[[na]] \notin [[varset]]$\\ Then
       $[[ [mu] (ord varset in iN) ]] = [[ [mu] (ord varset in na)]] = [[·]]$
       by \ruleref{\ottdruleOPVarNInLabel}.
       To prove that\\ $[[ ord [mu] varset in [mu] na ]] = [[·]]$, we apply
       \ruleref{\ottdruleOPVarNInLabel}. Let us show that
       $[[ [mu] na ]] \notin [[ [mu] varset ]]$.
       Since $[[ [mu] na ]] = [[mu]]([[na]])$ and
       $[[ [mu] varset ]] \subseteq [[mu]](A \cap [[varset]]) \cup [[varset]]$,
       it suffices to prove 
       $[[mu]]([[na]]) \notin [[mu]](A \cap [[varset]]) \cup [[varset]]$.

@@

  \item $[[iN]] = [[↑iP]]$
    \begin{align*}
       [[ [mu] (ord varset in iN) ]] &= [[ [mu] (ord varset in ↑iP) ]] \\
                                     &= [[ [mu] (ord varset in iP) ]]
                                     && \text{by \ruleref{\ottdruleOShiftULabel}}\\
                                     &= [[ ord [mu]varset in [mu]iP ]]
                                     && \text{by the induction hypothesis}\\
                                     &= [[ ord [mu]varset in  ↑[mu]iP ]]
                                     && \text{by \ruleref{\ottdruleOShiftULabel}}\\
                                     &= [[ ord [mu]varset in  [mu]↑iP ]]
                                     && \text{by the definition of substitution}\\
                                     &= [[ ord [mu]varset in  [mu]iN ]]
    \end{align*}

   \item $[[iN]] = [[iP → iM]]$
     \begin{align*}
        & [[ [mu] (ord varset in iN) ]] \\ 
        &= [[ [mu] (ord varset in iP → iM) ]] \\
        &= [[ [mu] (ordVars1, (ordVars2 \ {ordVars1})) ]]
          && \text{where } [[ord varset in iP = ordVars1]] \text{ and } [[ord varset in iM = ordVars2]] \\
        &= [[ [mu] ordVars1, [mu](ordVars2 \ {ordVars1}) ]] \\
        &= [[ [mu] ordVars1, ([mu]ordVars2 \ [mu]{ordVars1}) ]]
          && \text{by induction on $[[ordVars2]]$;
                  the inductive step is similar to 
                  \cref{case:distr-mu-ord:var}.}\\
        & && \text{Notice that $[[mu]]$ is collision free on $[[{ordVars1}]]$ and $[[{ordVars2}]]$} \\
        & && \text{since
          $[[{ordVars1}]] \subseteq [[varset]]$ and
          $[[{ordVars2}]] \subseteq [[fv iN]]$ }\\
          &= [[ [mu] ordVars1, ([mu]ordVars2 \ {[mu]ordVars1}) ]]
      \end{align*}
      \hfill\\
      On the other hand,
       $[[  ord [mu] varset in [mu]iN ]] = 
        [[ ord [mu] varset in [mu]iP → [mu]iM ]] = 
        [[ ordVarsb1, (ordVarsb2 \ {ordVarsb1}) ]] = 
        [[ [mu] ordVars1, ([mu]ordVars2 \ {[mu]ordVars1}) ]]$,
        where  $[[ord [mu] varset in [mu] iP = ordVarsb1]]$ 
        and $[[ord [mu] varset in [mu] iM = ordVarsb2]]$, 
        then by the induction hypothesis, 
        $[[ordVarsb1]] = [[ [mu] ordVars1 ]]$, 
        $[[ordVarsb2]] = [[ [mu] ordVars2 ]]$.

   \item $[[iN]] = [[∀ pas.iM]]$
     \begin{align*}
          [[ [mu] (ord varset in iN) ]] &= [[ [mu] ord varset in ∀pas.iM]] \\
                                        &= [[ [mu] ord varset in iM]] \\
                                        &= [[ ord [mu] varset in [mu] iM]]
                                        && \text {by the induction hypothesis}\\
     \end{align*}
     \begin{align*}
       [[ (ord [mu] varset in [mu] iN) ]] &= [[ ord [mu] varset in [mu] ∀pas.iM ]] \\
                                          &= [[ ord [mu] varset in ∀pas.[mu]iM ]] \\
                                          &= [[ ord [mu] varset in [mu] iM ]] \\
     \end{align*}
  \end{caseof}
\end{proof}

\lemOrdSigma*
\begin{proof}
  Mutual induction on $[[iN]]$ and $[[iP]]$.
  \begin{caseof}
    \item $[[iN = na]]$ \\
      If $[[na ∉ Γ1]]$ then $[[ [σ]na = na ]]$ and $[[ ord varset in [σ]na ]] = [[ ord varset in na ]]$, 
      as requried.
      If $[[na ∊ Γ1]]$ then $[[na ∉ varset]]$, so $[[ ord varset in na ]] = [[·]]$.
      Moreover, $[[Γ2 ⊢ σ : Γ1]]$ means $[[ fv([σ]na) ⊆ Γ2 ]]$, and thus, 
      as a set, $[[ ord varset in [σ]na ]] = [[varset ∩ fv([σ]na)]] \subseteq [[varset ∩ Γ2]] = [[·]]$.
    \item $[[iN = ∀pas.iM]]$
      We can assume $[[{pas} ∩ Γ1 = ∅]]$
      and $[[{pas} ∩ varset = ∅]]$. Then 
      \begin{align*}[t]
         [[ ord varset in [σ]iN ]] &= [[ ord varset in [σ]∀pas.iM ]] \\
                                   &= [[ ord varset in ∀pas.[σ]iM ]]\\
                                   &= [[ ord varset in [σ]iM ]]
                                   && \text{by the induction hypothesis}\\
                                   &= [[ ord varset in iM ]]\\
                                   &= [[ ord varset in ∀pas.iM ]]\\
                                   &= [[ ord varset in iN ]]
       \end{align*}
    \item $[[iN = ↑iP]]$
       \begin{align*}[t]
        [[ ord varset in [σ]iN ]] &= [[ ord varset in [σ]↑iP ]] \\
                                   &= [[ ord varset in ↑[σ]iP ]]
                                   && \text{by the definition of substitution}\\
                                   &= [[ ord varset in [σ]iP ]]
                                   && \text{by the induction hypothesis}\\
                                   &= [[ ord varset in iP ]]
                                   && \text{by the definition of substitution}\\
                                   &= [[ ord varset in ↑iP ]]
                                   && \text{by the definition of ordering}\\
                                   &= [[ ord varset in iN ]]
       \end{align*}

    \item $[[iN = iP → iM]]$
       \begin{align*}
        [[ ord varset in [σ]iN ]] &= [[ ord varset in [σ](iP → iM) ]] \\
                                   &= [[ ord varset in ([σ]iP → [σ]iM) ]]
                                   && \text{def. of substitution}\\
                                   &= [[ ord varset in [σ]iP]],\\
                                   &\phantom{=} ~ [[(ord varset in [σ]iM \ {ord varset in [σ]iP}) ]]
                                   && \text{def. of ordering}\\
                                   &= [[ ord varset in iP]],\\
                                   &\phantom{=} ~ [[(ord varset in iM \ {ord varset in iP}) ]]
                                   && \text{the induction hypothesis}\\
                                   &= [[ ord varset in iP → iM ]]
                                   && \text{def. of ordering}\\
                                   &= [[ ord varset in iN ]]
       \end{align*}
    \item The proofs of the positive cases are symmetric.
  \end{caseof}
\end{proof}
